Numerical modelling of the Cordilleran ice sheet

### 2014-08-25 kappa.tex
Add Swedish abstract.

% Document properties
\thesistitle{Numerical modelling of the Cordilleran ice sheet}
\runningtitle{Numerical modelling of the Cordilleran ice sheet}
\candidate{Julien Seguinot}
\runningauthor{J. Seguinot}
\pubyear{2014}
\department{Department of Physical Geography and Quaternary Geology}
\university{Stockholm University}
\issn{1653-7211}
\isbn{978-91-7447-973-7}
\coverpic{Numerical reconstruction of the Cordilleran ice sheet at the Last
          Glacial Maximum. Bedrock topography, ice surface topography, ice
          surface velocity, and frozen bed areas.}
\layout{Julien Seguinot based on {\LaTeX} templates by Rickard Petterson
        (summary) and Copernicus Publications (Paper~IV). Published papers
        typeset by Copernicus Publications (Paper~I) and the International
        Glaciology Society (Papers~II and~III). Graphics created using
        matplotlib, cartopy, {\LaTeX} and Inkscape.}
\printed{Universitetsservice US-AB, Stockholm, Sweden, 2014.}

@@

This doctoral dissertation presents a study of the glacial history of the North
American Cordillera using numerical ice sheet modelling calibrated against
field evidence. This area, characterized by the steep topography of several
mountain ranges separated by large inter-montane depressions, was once covered
by a large-scale ice mass: the former Cordilleran ice sheet. Because of the
irregular topography on which the ice sheet formed, geological studies have
often had only local or regional relevance, thus leaving the Cordilleran ice
sheet least understood among Pleistocene ice sheets in terms of its extent,
volume, and dynamics.

Here, I present numerical simulations that allow quantitative reconstructions
of the former ice
sheet evolution based on approximated physics of glacier flow. These
simulations show that the geometry of the Last Glacial Maximum Cordilleran ice
sheet was largely controlled by sharp contrasts in regional temperature,
precipitation, and daily temperature variability associated with the presence
of mountain ranges.

However, this maximum stage appears short-lived and out of balance with
contemporaneous climate. During most of the simulated last glacial cycle, the
North American Cordillera is characterized by an intermediate state of
glaciation including isolated glaciers and ice caps covering major mountain
ranges, the largest of which is located over the Skeena Mountains. The
numerically modelled Cordilleran ice sheet appears in constant imbalance with
evolving climate conditions, while the complexity of this transient response
transcends that encapsulated in two-dimensional, conceptual models of ice sheet
growth and decay.

@@

Hej!
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.
Här kommer sammanfattning på svenska. Här kommer sammanfattning på svenska.

>>>>

% Document properties
\thesistitle{Numerical modelling of the Cordilleran ice sheet}
\runningtitle{Numerical modelling of the Cordilleran ice sheet}
\candidate{Julien Seguinot}
\runningauthor{J. Seguinot}
\pubyear{2014}
\department{Department of Physical Geography and Quaternary Geology}
\university{Stockholm University}
\issn{1653-7211}
\isbn{978-91-7447-973-7}
\coverpic{Numerical reconstruction of the Cordilleran ice sheet at the Last
          Glacial Maximum. Bedrock topography, ice surface topography, ice
          surface velocity, and frozen bed areas.}
\layout{Julien Seguinot based on {\LaTeX} templates by Rickard Petterson (cover
        summary) and Copernicus Publications (Paper~IV). Published papers
        typeset by Copernicus Publications (Paper~I) and the International
        Glaciology Society (Papers~II and~III). Graphics created using
        matplotlib, cartopy, {\LaTeX} and Inkscape.}
\printed{Universitetsservice US-AB, Stockholm, Sweden, 2014.}

@@

\noindent
This doctoral dissertation presents a study of the glacial history of the North
American Cordillera using numerical ice sheet modelling calibrated against
field evidence. This area, characterized by the steep topography of several
mountain ranges separated by large inter-montane depressions, was once covered
by a large-scale ice mass: the former Cordilleran ice sheet. Because of the
irregular topography on which the ice sheet formed, geological studies have
often had only local or regional relevance, thus leaving the Cordilleran ice
sheet least understood among Pleistocene ice sheets in terms of its extent,
volume, and dynamics.

Here, I present numerical simulations that allow quantitative reconstructions
of the former ice
sheet evolution based on approximated physics of glacier flow. These
simulations show that the geometry of the Last Glacial Maximum Cordilleran ice
sheet was largely controlled by sharp contrasts in regional temperature,
precipitation, and daily temperature variability associated with the presence
of mountain ranges.
However, this maximum stage appears short-lived and out of balance with
contemporaneous climate. During most of the simulated last glacial cycle, the
North American Cordillera is characterized by an intermediate state of
glaciation including isolated glaciers and ice caps covering major mountain
ranges, the largest of which is located over the Skeena Mountains. The
numerically modelled Cordilleran ice sheet appears in constant imbalance with
evolving climate conditions, while the complexity of this transient response
transcends that encapsulated in two-dimensional, conceptual models of ice sheet
growth and decay.

@@

\noindent
I min avhandling presenterar jag den glaciala historien för de Nordamerikanska
Kordillerabergen baserad på numerisk ismodellering kalibrerad mot fältdata.
Området, som karaktäriseras av branta bergskedjor och mellanliggande lägre
platåer, täcktes tidigare av en inlandsis, Kordilleraisen. Den oregelbundna
topografin över vilken inlandsisen bildades har bidragit till att tidigare
fältstudier främst har haft ett lokalt eller regionalt fokus och bland de
Pleistocena istäckena hör därför Kordilleraisen till de sämst utredda när det
gäller inlandsisens utbredning, volym och dynamik.

Jag presenterar numeriska simuleringar som är baserade på approximationer av
isflödesfysik och som möjliggör kvantitativa rekonstruktioner av istäckets
utveckling. Simuleringarna visar att Kordilleraisen under dess senaste glaciala
maxima till stor del styrdes av markanta regionala kontraster i temperatur- och
nederbördsförhållanden samt dygnsmässiga temperaturvariationer.
Inlandsisens största utbredning under den sista glaciala cykeln var kortlivad.
Under största delen av den simulerade glaciala cykeln karaktäriserades
nedisningsområdet av isfält och platåisar samt isolerade glaciärer i
bergskedjorna, och störst isvolym fanns över Skeenafjället. Den modellerade
Kordilleraisen verkar ha varit i konstant obalans med det ständigt förändrande
klimatet. Inlandsisens respons på dessa klimatförändringar var därför vidare
mer komplext än vad som har presenterats i tvådimensionella konceptuella
modeller av hur inlandsisen växte till och smälte bort.

I min avhandling demonstrerar jag den potential som numerisk ismodellering har
för att kunna dra meningsfulla slutsatser om nedisningshistoriken och forna
klimatförhållanden över glaciala cykler, under förutsättning att modelleringen
kan kalibreras mot fältdata.
